---
title: "Tennis Grand Slams"
author: 'null'
date: '2020-12-08'
slug: tennis-grand-slams
categories: []
tags: []
---

# Men's and Women's Tennis Grand Slam Championships

```{r, error=FALSE, message=FALSE, warning=FALSE}
library(here)
library(tidyverse)
library(ggplot2)

dat1 <- read_csv(here::here("data", "grand_slam_timeline.csv"))
  
dat2 <- read_csv(here::here("data", "grand_slams.csv"))
  
dat3 <- read_csv(here::here("data", "player_dob.csv"))


```

The first data set comes from the `grand_slam_timeline.csv` file on the Tidy Tuesday website. The data set has 12,605 observations of 5 variables. The variables are fairly self-explanatory, `player` is the name of the player, `year` is the year of the tournament, `tournament` gives the name of the tournament, `outcome` gives what position in the tournament the player finished, and `gender` gives the gender of the player. 

The second data set comes from the `grand_slams.csv` file on the Tidy Tuesday website. This data set has only 416 observations on 6 variables. This data set provides the same variables `year` and `gender` as the previous data set. Similarly, `grand_slam` provides the same information as `tournament` in the previous data set, and `name` provides the same information as `player`. This data set additionally provides the `rolling_win_count` which gives the total number of wins for the player at the time of that tournament, and `tournament_date` which gives the approximate date that the tournament took place. 

The third data set comes from the `player_dob.csv` file on the Tidy Tuesday website. This data set has 105 observations of 5 variables. The `name` variable is the same as in the previous data set. For this data set, `grand_slam` indicates what tournament the player was playing at when they won their first grand slam. `date_of_first_title` gives the date at which this occurs. `date_of_birth` gives the date of birth of the player, allowing `age` to be calculated by finding the difference in number of days between the `date_of_first_title` and `date_of_birth` to give the age at the time of first grand slam championship. 

## Question 1
What grand slam is most common to win a first title at?
  * I will explore this by using the third data set to create a bar chart displaying how many players won their first grand slam at a given tournament. This could also be explored with facets such as age and gender. 

```{r, warning=FALSE, message=FALSE, error=FALSE}
dat3 %>%
  filter(!is.na(grand_slam)) %>%
  ggplot(., aes(x = grand_slam)) +
  geom_bar() +
  ggtitle("Tournament at Which First Grand Slam Title Was Won") +
  ylab("Count of People") +
  xlab("Tournament")

```

## Question 2
Do players that are younger at the time of first grand slam win more tournaments overall?
  * I will analyze this by plotting the age at first win vs. number of rolling wins at the end of the data set. This could also be analyzed broken down into gender categories. 
  
```{r, error=FALSE, message=FALSE, warning=FALSE}

dat2 %>%
  group_by(name) %>%
  filter(rolling_win_count == max(rolling_win_count)) %>%
  left_join(dat3, c("name" = "name")) %>%
  filter(!is.na(age)) %>%
  mutate(age_years = (age/365.25)) %>%
  ggplot(., aes(x = age_years, y = rolling_win_count)) +
  geom_point() +
  geom_smooth() +
  ylim(0, 25) +
  ggtitle("Total Number of Grand Slams Won vs. Age at First Win") +
  xlab("Age (in years)") +
  ylab("Total number of wins")

dat2 %>%
  group_by(name) %>%
  filter(rolling_win_count == max(rolling_win_count)) %>%
  left_join(dat3, c("name" = "name")) %>%
  filter(!is.na(age)) %>%
  mutate(age_years = (age/365.25)) %>%
  ggplot(., aes(x = age_years, y = rolling_win_count, color = gender)) +
  geom_point() +
  geom_smooth() +
  ylim(0, 25) +
  ggtitle("Total Number of Grand Slams Won vs. Age at First Win by Gender") +
  xlab("Age (in years)") +
  ylab("Total number of wins")
```
  
## Question 3
How does the age of first title won differ between males and females?
  * This can be analyzed by creating a histogram of the age at which the first title is won and faceting by gender. 

```{r, warning=FALSE, message=FALSE, error=FALSE}
dat2 %>%
  group_by(name) %>%
  filter(rolling_win_count == max(rolling_win_count)) %>%
  left_join(dat3, c("name" = "name")) %>%
  filter(!is.na(age)) %>%
  mutate(age_years = (age/365.25)) %>%
  ggplot(., aes(x = age_years)) +
  geom_histogram() +
  facet_wrap("gender") +
  ggtitle("Distribution of Age at Which First Grand Slam was Won by Gender") +
  xlab("Age (in years)") +
  ylab("Number of people")

```


## Question 4
What tournament was most frequently missed? 
  * This can be found by creating a bar chart of the tournaments and the counts of the number of players that were absent from the tournament. Additionally this could be split into male and female categories.
  
```{r, warning=FALSE, error=FALSE, message=FALSE}
dat1 %>%
  filter(outcome == "Absent") %>%
  ggplot(., aes(x = tournament)) +
  geom_bar() +
  ggtitle("Count of the Number of People That Missed Each Grand Slam") +
  xlab("Tournament") +
  ylab("Count of players")

dat1 %>%
  filter(outcome == "Absent") %>%
  ggplot(., aes(x = tournament)) +
  geom_bar() +
  facet_wrap("gender") +
  ggtitle("Count of the Number of People That Missed Each Grand Slam by Gender") +
  xlab("Tournament") +
  ylab("Count of players")
```